---
title: "Topic modeling - Svevo corpus of italian letters"
output: html_document
---

# LDA 

Latent Dirichlet Allocation has been use to model topics in Svevo Corpus of italian letters.
The main parameter to be tuned in order to use LDA is the number of topics k and to obtain a suitable 
value of k we started training models on the corpus of italian letters for k in [2,50]. 
Then, for all the trained models three indexes are compared: perplexity, coherence of topics and silhouette.
By combining all the previous metrics, a suitable value for k is chosen.

  
```{r load dataset, warning = FALSE}
# load functions for LDA analysis
source("LDA/lda_functions.R")
library(dplyr)
library(tidytext)
# Load dataset with lemmatized words
corpus_ita <- read.csv("csv/cleaned_svevo_dataset_ITA.csv", sep=",", encoding = "UTF-8")


```

```{r evaluate perplexity and coherence, eval=FALSE, warning = FALSE}
# take a lot of time !!!
perplexity_lda <- evaluate_perplexity(corpus_ita, save_results = TRUE)

coherence_lda <- evaluate_coherence(kmax = 50, corpus_ita, save_results = TRUE)

```

```{r plot perplexity and coherence}
perplexity_lda <- read.csv("csv/perplexity.csv")

ggplot(perplexity_lda, aes(x = k, y = perplexity)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  ggtitle("5-fold cross-validation of topic modelling",
          "(The points represent five different models fit for each candidate number of topics)") +
  labs(x = "K", y = "Perplexity when fitting the trained model to the hold-out set")

#ggsave("plots/perplexity.png", width = 20, height = 8, dpi = 150)

coherence_lda <- read.csv("csv/coherence.csv")
colnames(coherence_lda) <- c("k", "coherence")

ggplot(coherence_lda) +
  geom_point(aes(x = 5, y = coherence_lda$coherence[5]), col = "red", size = 3) +
  geom_line(aes(x = k, y = coherence), col = "violet") + 
  xlab("K") + 
  ylab("Coherence")

#ggsave("plots/coherence.png", width = 20, height = 8, dpi = 150)

```
```{r plot silhouette}
silhouette_lda <- read.csv("csv/silhouette.csv")

ggplot(silhouette_lda, aes(x = X, y = x)) +
  geom_point() +
  geom_point(aes(x = 5, y = silhouette_lda$x[5]), col = "red") +
  geom_smooth(se = FALSE) +
  labs(x = "K", y = "silhouette score")

#ggsave("plots/silhouette.png", width = 20, height = 8, dpi = 150)
```




# Main Topics analysis

We choose an intermediate value for k, which ensures a good coherence between topics
while (keeping perplexity low) ensuring a good silhouette score.

Build LDA model with a suitable value for k.
```{r build lda model, eval=FALSE, include=FALSE, warning = FALSE}
lda_model_ita <- one_model_analysis(num_topics = 5, corpus_ita, save_results = TRUE)
```

Load pre-trained LDA model

```{r }
lda_model_ita <- readRDS("LDA/LDA_corpus_topic_model.rds")
```

Analysis of main topics obtained.
```{r}
#print summary table
lda_model_ita$summary[ order(lda_model_ita$summary$prevalence, decreasing = TRUE) , ]
SummarizeTopics(lda_model_ita)

```


Now assign suitable labels to topics:

```{r}

lda_model_ita$labels <- c("salute", "famiglia", "libro", "pensieri", "viaggio" )

```

Analyze how topics evolved over time:

```{r}
#input: one topic
#output: its trend over time

# obtain document-topic probabilities -> lda_ita$theta
# number are those of italian letters in the original corpus

topic_time <- topic_trend_over_time(corpus_ita, lda_model_ita)
names(topic_time)[1:5] <- lda_model_ita$labels

```

```{r plot topic trend over time}
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape2)

sum_topic <- topic_time %>%
  group_by(date) %>%
  summarise(across(everything(), sum)) 

count <- (topic_time %>% count(date))$n
sum_topic$n <- count


sum_topic %>%
    gather(key, value, lda_model_ita$labels) %>%
    ggplot() +
    geom_point(aes(x=date, y=value/n, colour=key, group = 1),ylab="") +
    geom_line(aes(x=date, y=value[1:30]/n, colour=key[1], group = 1),ylab="")+
    geom_line(aes(x=date, y=value[31:60]/n, colour=key[31], group = 1),ylab="")+
    geom_line(aes(x=date, y=value[61:90]/n, colour=key[61], group = 1),ylab="")+
    geom_line(aes(x=date, y=value[91:120]/n, colour=key[91], group = 1),ylab="")+
    geom_line(aes(x=date, y=value[121:150]/n, colour=key[121], group = 1),ylab="")

sum_topic <- melt(sum_topic, id = c("date","n"))

sum_topic %>%
  ggplot() + 
  geom_bar(aes(x = date, y = value/n, fill = variable), position = "dodge", stat = "identity")
```

```{r}
#input: one topic
#output: to who the topic is assigned to

# obtain document-topic probabilities -> lda_ita$theta
# number are those of italian letters in the original corpus

topic_people <- topic_trend_over_people(corpus_ita, lda_model_ita)
names(topic_people)[1:5] <- lda_model_ita$labels

```

```{r plot topic trend over person}
library(dplyr)
library(ggplot2)
library(tidyr)

sum_topic <- topic_people %>%
  group_by(pair) %>%
  mutate(count = n()) %>%
  filter(count > 4) %>%
  group_by(pair, count) %>%
  summarise(across(everything(), sum)) 


sum_topic <- melt(sum_topic, id = c("pair", "count"))

sum_topic %>%
  ggplot() + 
  geom_bar(aes(x = pair, y = value/count, fill = variable), position = "dodge", stat = "identity")

```


